Implementing Equational Constraints\\ in a Functional Language

### 2014-04-08 docs/src/currytest.tex

We can execute a test suite by the command\pindex{currytest}
\begin{curry}
currytest TestList
\end{curry}
(\code{currytest} is a program stored in \code{\cyshome/bin}
where \cyshome is the installation directory of \CYS;
see Section~\ref{sec-install}).
In our example, \ccode{TestList.curry} is the program containing the
definition of all assertions. This has the effect
that all exported top-level functions
of type \code{Assertion} are tested (i.e., the corresponding
assertions are checked) and the results
(\ccode{OK} or failure) are reported together with the name of each assertion.
%If failures occur, the complete test results are also
%written into a file named \ccode{TestList.testlog}.''
For our example above, we obtain the following successful protocol:
\begin{curry}
============================================================
Testing module "TestList"...
OK: ++
OK: all
OK: prefix
All tests successfully passed.
============================================================
\end{curry}
There is also a graphical interface that summarizes the results
more nicely.
In order to start this interface, one has to add the parameter
\ccode{--window} (or \ccode{-w}), e.g., executing a test suite by
\begin{curry}
currytest --window TestList
\end{curry}
or
\begin{curry}
currytest -w TestList
\end{curry}
A snapshot of the interface is shown in Figure~\ref{fig-currytest}.

>>>>

We can execute a test suite by the command\pindex{currytest}
\begin{curry}
currytest TestList
\end{curry}
(\code{currytest} is a program stored in \code{\cyshome/bin}
where \cyshome is the installation directory of \CYS;
see Section~\ref{sec-general}).
In our example, \ccode{TestList.curry} is the program containing the
definition of all assertions. This has the effect
that all exported top-level functions
of type \code{Assertion} are tested (i.e., the corresponding
assertions are checked) and the results
(\ccode{OK} or failure) are reported together with the name of each assertion.
%If failures occur, the complete test results are also
%written into a file named \ccode{TestList.testlog}.''
For our example above, we obtain the following successful protocol:
\begin{curry}
============================================================
Testing module "TestList"...
OK: ++
OK: all
OK: prefix
All tests successfully passed.
============================================================
\end{curry}
There is also a graphical interface that summarizes the results
more nicely.
In order to start this interface, one has to add the parameter
\ccode{--window} (or \ccode{-w}), e.g., executing a test suite by
\begin{curry}
currytest --window TestList
\end{curry}
or
\begin{curry}
currytest -w TestList
\end{curry}
A snapshot of the interface is shown in Figure~\ref{fig-currytest}.

### 2014-04-08 docs/src/usage.tex

\item[\fbox{\code{+/-safe}}]\pindex{safe}
Turn on/off the safe execution mode.
In the safe execution mode, the initial goal is
not allowed to be of type \code{IO} and the program should not
import the module \code{Unsafe}.
Furthermore, the allowed commands are
\code{eval}, \code{load}, \code{quit}, and \code{reload}.
This mode is useful to use \CYS in uncontrolled environments,
like a computation service in a web page.

>>>>

\item[\fbox{\code{safe}}]\pindex{safe}
Turn on the safe execution mode.
In the safe execution mode, the initial goal is
not allowed to be of type \code{IO} and the program should not
import the module \code{Unsafe}.
Furthermore, the allowed commands are
\code{eval}, \code{load}, \code{quit}, and \code{reload}.
This mode is useful to use \CYS in uncontrolled environments,
like a computation service in a web page, where \CYS could
be invoked by
\begin{curry}
kics2 :set safe
\end{curry}

### 2014-04-08 talks/2013-padl/slides.tex

\date{WLP 2011}
\author[Michael Hanus]{Michael Hanus\\[2ex]
{\small Joint work with Bernd Bra\ss{}el, Bj\"orn Peem\"oller and Fabian Reck}}
\institute[CAU Kiel]{Christian-Albrechts-University of Kiel\\
  Programming Languages and Compiler Construction}

@@

\begin{block}{Goal: combine best of declarative paradigms in a single model}
\begin{itemize}%\expandlist{1ex}
\item {\blue efficient execution} principles of functional languages\\
(determinism, laziness)
\item {\blue flexibility} of logic languages\\
(computation with partial information, built-in search)
\item {\blue application-domains} of constraint languages\\
(constraint solvers for specific domains)
\item avoid non-declarative features of Prolog\\
(arithmetic, cut, I/O, side-effects)
\end{itemize}
\end{block}
\vfill
\pause

\begin{block}{Curry {\lgrey [POPL'97,\ldots]} $\leadsto$ {\tt http://www.curry-language.org/}}
\begin{itemize}%\expandlist{1ex}
\item declarative multi-paradigm language\\
 (higher-order concurrent functional logic language)
\item extension of Haskell (non-strict functional language)
\item SWE advantage: better (high-level) APIs for various application domains\\
(GUI programming, web programming, database programming,\ldots)
\end{itemize}
\end{block}
\end{frame}

@@

\begin{exampleblock}{Example \hspace{4cm}\only<3>{(with \alert{equational constraint}~ \texttt{=:=})}\only<4>{(with \alert{functional pattern})}}
\begin{curry}
(++) :: [a] -> [a] -> [a]    $\only<3-4>{\code{last :: [a] -> a}}$
[]     ++ ys = ys              $\only<3>{\code{last xs | \alert{ys ++ [x] =:= xs}}}$$\only<4>{\code{last \alert{(ys ++ [x])} = x}}$
(x:xs) ++ ys = x : xs ++ ys             $\only<3>{\code{= x  where ys,x free}}$
\end{curry}
\end{exampleblock}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
\frametitle{Implementing Curry}

\begin{block}{Abstract machines}
\begin{itemize}
\item C (MCC, M\"unster Curry Compiler)
\item Java (FLVM, Portland)
\end{itemize}
\end{block}
\pause

\begin{block}{Compiling to Prolog}
\begin{itemize}
\item PAKCS (Portland Aachen Kiel Curry System), TOY (Madrid)
\item reuse non-determinism (backtracking), logic variables, constraint solvers
\item implement laziness, higher-order
\end{itemize}
\end{block}
\pause

\begin{block}{Compiling to Haskell: {\bf KiCS2} {\lgrey[WFLP'11]}}
\begin{itemize}
\item reuse laziness, higher-order
\item implement non-determinism, logic variables
\item advantages: \alert{search strategy not fixed}\\
      (backtracking, breadth-first, iterative deepening, parallel,$\ldots$)
\end{itemize}
\end{block}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
\frametitle{Compiling Curry to Haskell: KiCS2}

\begin{block}{Principles of KiCS2}
\begin{itemize}\expandlist{2ex}
\item represent search space as data ($\approx$ search tree)
\item defined operation: generate \emph{all} its values
\item search strategy $\approx$ search tree traversal
\item Haskell's laziness: no problem with infinite trees
\end{itemize}
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
\frametitle{Representing Non-deterministic Values}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
\frametitle{Depth-first Search}
\vspace{-1ex}
\begin{exampleblock}{Print all values of a choice structure in depth-first order}
\vspace{-1ex}
\begin{curry}
  printValsDFS :: a -> IO ()$\listline$
  printValsDFS Fail             = return ()$\listline$
  printValsDFS (Choice i x1 x2) = lookupDecision i >>= follow
   where
    follow ChooseLeft  = printValsDFS x1
    follow ChooseRight = printValsDFS x2
    follow NoDecision    = do newDecision ChooseLeft  x1
                              newDecision ChooseRight x2

    newDecision d x = do setDecision i d
                         printValsDFS x
                         setDecision i NoDecision$\listline$
  printValsDFS v = print v
\end{curry}
\vfill\pause

Main call: \code{initSupply >>= \bs{}s -> printValsDFS (main s)}\\
Result:~~~~ \code{False}\\
\hspace{1.62cm}\code{False}
\end{exampleblock}

@@

\begin{block}{}
\citebib{Antoy/H. ICLP'06}:\\[1ex]
narrowing (resolution) on logic variables \\
$~~~~~~~~~~~~~~~~~~~~~~~\Updownarrow$\\
non-deterministic rewriting with generators
\end{block}
\end{frame}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{frame}
\frametitle{Conclusions}
\vspace{-1ex}
\begin{block}{}
\begin{itemize}
\item novel implementation techniques for equational constraints
\item add bindings constraints to computed values
\item process them when extracting values from search space
\item only new constructors added, no overhead for existing implementation
\item more efficient execution if features are used
\end{itemize}
\end{block}
\vfill
\begin{block}{Advantages of KiCS2}
\begin{itemize}
\item highly efficient for deterministic parts, competitive for non-det. programs
\item flexible search strategies
\item many opportunities for optimizations (determinism analysis)
\item exploit future improvements of Haskell implementations
\end{itemize}
\end{block}
\vfill
\begin{block}{Future Work}
\begin{itemize}
\item Add further constraint structures
\end{itemize}
\end{block}
\end{frame}

>>>>

\date{PADL 2013}
\author[Fabian Reck]{Fabian Reck\\[2ex]
{\small Joint work with Bernd Bra\ss{}el, Michael Hanus and Bj\"orn Peem\"oller}}
\institute[CAU Kiel]{Christian-Albrechts-University of Kiel\\
  Programming Languages and Compiler Construction}

@@

\begin{block}{Goal: combine best of declarative paradigms in a single model}
\begin{itemize}%\expandlist{1ex}
\item {\blue efficient execution} principles of functional languages\\
(determinism, laziness)
\item {\blue flexibility} of logic languages\\
(computation with partial information, built-in search)
\item {\blue application-domains} of constraint languages\\
(constraint solvers for specific domains)
\item avoid non-declarative features
\end{itemize}
\end{block}
\vfill
\pause

\begin{block}{Curry {\lgrey [POPL'97,\ldots]} $\leadsto$ {\tt http://www.curry-language.org/}}
\begin{itemize}%\expandlist{1ex}
\item declarative multi-paradigm language\\
 (higher-order concurrent functional logic language)
\item extension of Haskell (non-strict functional language)
\end{itemize}
\end{block}
\end{frame}

@@

\begin{exampleblock}{Example \hspace{4cm}\only<3>{(with \alert{equational constraint}~ \texttt{=:=})}}
\begin{curry}
(++) :: [a] -> [a] -> [a]    $\only<3>{\code{last :: [a] -> a}}$
[]     ++ ys = ys              $\only<3>{\code{last xs | \alert{ys ++ [x] =:= xs}}}$
(x:xs) ++ ys = x : xs ++ ys             $\only<3>{\code{= x  where ys,x free}}$
\end{curry}
\end{exampleblock}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
\frametitle{Implementing Curry: KiCS2}

\begin{block}{Compiling to Haskell: {\bf KiCS2} {\lgrey[WFLP'11]}}
\begin{itemize}
\item reuse laziness, higher-order
\item implement non-determinism, logic variables
\item advantages: \alert{search strategy not fixed}\\
      (backtracking, breadth-first, iterative deepening, parallel,$\ldots$)
\end{itemize}
\end{block}
\pause
\begin{block}{Principles of KiCS2}
\begin{itemize}\expandlist{2ex}
\item represent search space as data ($\approx$ search tree)
\item defined operation: generate \emph{all} its values
\item search strategy $\approx$ search tree traversal
\item Haskell's laziness: no problem with infinite trees
\end{itemize}
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
\frametitle{Representing Non-deterministic Results}

@@

@@

\begin{block}{}
\citebib{Antoy/Hanus ICLP'06}:\\[1ex]
narrowing (resolution) on logic variables \\
$~~~~~~~~~~~~~~~~~~~~~~~\Updownarrow$\\
non-deterministic rewriting with generators
\end{block}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{frame}
\frametitle{Conclusions}
\vspace{-1ex}
\begin{block}{}
\begin{itemize}
\item novel implementation techniques for equational constraints
\item add binding constraints to computed values
\item process them when extracting values from search space
\item only new constructors added, no overhead for existing implementation
\item more efficient execution if features are used
\end{itemize}
\end{block}
\vfill
\begin{block}{Advantages of KiCS2}
\begin{itemize}
\item highly efficient for deterministic parts, competitive for non-det. programs
\item flexible search strategies
\item many opportunities for optimizations (determinism analysis)
\item exploit future improvements of Haskell implementations
\end{itemize}
\end{block}
\vfill
\begin{block}{Future Work}
\begin{itemize}
\item Add further constraint structures
\end{itemize}
\end{block}
\end{frame}

@@
